From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-08 Pi+/paper/ufin.tex
Switch theorems to theoremreps in sections 3 and 4

\begin{proposition}[Univalent Subuniverse]
  \label{prop:univsub}
  A universe predicate is a type family $P : \UU \to \UU$ whose fibers are propositions, that is, $P(X)$ is a
  proposition for every $X$. Given such a predicate, the fibration ${\fst : \dsum*{X:\UU}{P(X)} \to \UU}$ is univalent and generates a
  univalent subuniverse $\UU_{P} \defeq (\dsum{X:\UU}{P(X)}, \fst)$.
\end{proposition}

\begin{appendixproof}
  (of Prop.~\ref{prop:univsub})
  Suppose $(U, \El) \defeq (\dsum*{X:\UU}{P(X)}, \fst)$ is a subuniverse generated by a subtype $P : \UU \to \UU$. For
  any $X, Y : \UU$ such that $\phi : P(X)$ and $\psi : P(Y)$, we want to show that
  $\tptEqv{\fst} : (X,\phi) \id (Y,\psi) \to X \eqv Y$ is an equivalence. We construct
  $X \eqv Y \to (X,\phi) \id (Y,\psi)$ by $\ua$ and using the fact that $P(\blank)$ is a proposition. That it is an
  inverse follows by calculation using the appropriate computation rules.
\end{appendixproof}

@@

\begin{proposition}
  \label{prop:isFin}
  For any type $X$, $\isFin[X]$ is a proposition.
\end{proposition}

\begin{appendixproof}
  (of Prop.~\ref{prop.isFin})
  Suppose we have $(n,\phi) : \isFin[X]$ and $(m,\psi) : \isFin[X]$, we need to show that $(n,\phi) \id (m,\psi)$. It is
  enough to show that $n \id m$. Since $\Nat$ is a set, this is a proposition, so we can use the induction principle of
  propositional truncation to eliminate to $n \id m$, applying it on $\phi$ and $\psi$ respectively. This gives us the
  equalities $X \id \Fin[n]$ and $X \id \Fin[m]$, which gives us $\Fin[n] \id \Fin[m]$, from which $n \id m$ follows by
  applying the first projection.
\end{appendixproof}

@@

\begin{proprep}
  \leavevmode
  \begin{enumerate}
    \item If $T$ is an $n$-type, $\BAut[T]$ is an $(n+1)$-type.
    \item For any $T : \UU$, $\BAut[T]$ is 0-connected.
    \item For any $T : \UU$, \( \loopspace[\BAut[T],T_{0}] \eqv \Aut[T] \). \label{lem:loop-deloop}
  \end{enumerate}
\end{proprep}
\begin{proofsketch}
  See Appendix.
\end{proofsketch}

\begin{appendixproof}
  We need to show that the equality type of $\BAut[T]$ is an $n$-type. Assume $X, Y : \BAut[T]$. Since $\BAut[T]$ is a
  univalent subuniverse, we have $(X \id Y) \eqv (\fst(X) \eqv \fst(Y))$. Note that being an $n$-type is a proposition.
  Since $T$ is an $n$-type, and $\fst(X)$ and $\fst(Y)$ are merely equal to $T$, they're also $n$-types. It follows that
  $\fst(X) \eqv \fst(Y)$ is an $n$-type, and hence $X \id Y$ is an $n$-type.

  Since $\BAut[T]$ is a univalent universe, it follows that
  \[
    (T_{0} \id_{\BAut[T]} T_{0}) \eqv (\fst(T_{0}) \eqv \fst(T_{0})) \equiv (T \eqv T) \equiv \Aut[T].
  \]
\end{appendixproof}

>>>>

\begin{propositionrep}[Univalent Subuniverse]
  \label{prop:univsub}
  A universe predicate is a type family $P : \UU \to \UU$ whose fibers are propositions, that is, $P(X)$ is a
  proposition for every $X$. Given such a predicate, the fibration ${\fst : \dsum*{X:\UU}{P(X)} \to \UU}$ is univalent and generates a
  univalent subuniverse $\UU_{P} \defeq (\dsum{X:\UU}{P(X)}, \fst)$.
\end{propositionrep}

\begin{proof}
  Suppose $(U, \El) \defeq (\dsum*{X:\UU}{P(X)}, \fst)$ is a subuniverse generated by a subtype $P : \UU \to \UU$. For
  any $X, Y : \UU$ such that $\phi : P(X)$ and $\psi : P(Y)$, we want to show that $\tptEqv{\fst} : (X,\phi) \id
    (Y,\psi) \to X \eqv Y$ is an equivalence. We construct $X \eqv Y \to (X,\phi) \id (Y,\psi)$ by $\ua$ and using the
  fact that $P(\blank)$ is a proposition. That it is an inverse follows by calculation using the appropriate computation
  rules.
\end{proof}

@@

\begin{propositionrep}
\label{prop:isFin}
For any type $X$, $\isFin[X]$ is a proposition.
\end{propositionrep}

\begin{proof}
  (of Prop.~\ref{prop.isFin})
  Suppose we have $(n,\phi) : \isFin[X]$ and $(m,\psi) : \isFin[X]$, we need to show that $(n,\phi) \id (m,\psi)$. It is
  enough to show that $n \id m$. Since $\Nat$ is a set, this is a proposition, so we can use the induction principle of
  propositional truncation to eliminate to $n \id m$, applying it on $\phi$ and $\psi$ respectively. This gives us the
  equalities $X \id \Fin[n]$ and $X \id \Fin[m]$, which gives us $\Fin[n] \id \Fin[m]$, from which $n \id m$ follows by
  applying the first projection.
\end{proof}

@@

\begin{propositionrep}
  \leavevmode
  \begin{enumerate}
    \item If $T$ is an $n$-type, $\BAut[T]$ is an $(n+1)$-type.
    \item For any $T : \UU$, $\BAut[T]$ is 0-connected.
    \item For any $T : \UU$, \( \loopspace[\BAut[T],T_{0}] \eqv \Aut[T] \). \label{lem:loop-deloop}
  \end{enumerate}
\end{propositionrep}

\begin{proof}
  We need to show that the equality type of $\BAut[T]$ is an $n$-type. Assume $X, Y : \BAut[T]$. Since $\BAut[T]$ is a
  univalent subuniverse, we have $(X \id Y) \eqv (\fst(X) \eqv \fst(Y))$. Note that being an $n$-type is a proposition.
  Since $T$ is an $n$-type, and $\fst(X)$ and $\fst(Y)$ are merely equal to $T$, they're also $n$-types. It follows that
  $\fst(X) \eqv \fst(Y)$ is an $n$-type, and hence $X \id Y$ is an $n$-type.

  Since $\BAut[T]$ is a univalent universe, it follows that
  \[
    (T_{0} \id_{\BAut[T]} T_{0}) \eqv (\fst(T_{0}) \eqv \fst(T_{0})) \equiv (T \eqv T) \equiv \Aut[T].
  \]
\end{proof}
